Towards Proof-Reconstruction of Problems in Classical Propositional Logic

### 2017-10-04 paper/sections/metis-language-and-proofs.tex

\begin{figure}
\centering
\begin{bprooftree}\tt
  \AxiomC{}
  \RightLabel{\footnotesize negate}
  \UnaryInfC{$\neg p$}
  \RightLabel{\footnotesize strip}
  \UnaryInfC{$\neg p$}
  \AxiomC{}
  \RightLabel{\footnotesize axiom}
  \UnaryInfC{$p$}
  \RightLabel{\footnotesize canonicalize}
  \UnaryInfC{$p$}
  \RightLabel{\footnotesize simplify}
  \BinaryInfC{$⊥$}
  \RightLabel{\footnotesize canonicalize}
  \UnaryInfC{$⊥$}
\end{bprooftree}
\caption{The \Metis' proof tree for $p \vdash p$ from the
derivation in Fig.~\ref{fig:metis-proof-tstp}}
\label{fig:metis-example}
\end{figure}

@@

\textit{Splitting}. To prove a goal, \Metis splits the goal into
disjoint cases. This process produces a list of new subgoals, the
conjunction of these subgoals implies the goal as we show later in
Subsection~\ref{ssec:emulating-inferences}. Then, a proof of the
goal becomes in smaller proofs, one refutation for each subgoal.
These subgoals are introduced in the \TSTP derivation with the \strip
inference rule.

@@

\textit{Clausification.} The \clausify rule transforms a
propositional formula in clausal normal form or \CNF, a conjunction
of clauses. Where a clause is the disjunction of zero or more
literals and a literal is an atom (positive literal) or a negation
of an atom (negative literal). This conversion is not unique, and
\Metis performs this transformation using its criteria.
\improvement{We should mention that it was indeed a problem for the
proof-reconstruction? and refer the section next?}

@@

\textit{Normalization.} \Metis uses the \canonicalize rule to
normalize a propositional formula that comes from an axiom or a
definition. It often converts each expression in one of its normal
form: the conjunctive normal form, the negative normal form or the
disjunctive normal form. Next, this rule simplifies the formula
with some definitions, applying recursively on the structure's
formula conjunctions and disjunctions simplifications to remove
tautologies (see Fig.~\ref{fig:conjunctive-disjunctive-simpl} for a
list of these theorems. We assume the commutative property for the
conjunction and disjunction connectives).

@@

\textit{Resolution.} The \resolve rule is the resolution theorem
showed in Fig.~\ref{fig:metis-inferences}. For its application, we
need the positive literal for resolution and two derivations to
deduce the \emph{resolvent}. The positive literal $p$ must occur in
the formula from the first derivation, and the negative literal must
occur in the formula from the second derivation (this follows the
same pattern in \Metis' logic core in Fig.~\ref{fig:metis-inferences}
for the resolve inference).

@@

\textit{Splitting a conjunct.} The \conjunct rule extracts from a
conjunction one of its conjuncts; it is a generalization of the
projection rules for the conjunction connective as the
following \TSTP derivation shows.

@@

\textit{Negate.}
Each subgoal proof is a refutation, thereby each proof assumes
the negation of its subgoal. The \negate rule
introduces the negation of a subgoal that results
after applies the \strip inference to the goal.

@@

\textit{Simplification.}
The \simplify rule is an inference that performs simplification of
definitions and tautologies. This rule
transverses a list of formulas by applying different theorems (e.g.,
\emph{modus pones}, \emph{modus tollens}, or \emph{disjunctive
syllogism}) including the list of theorems used by \canonicalize in
Fig. \ref{fig:conjunctive-disjunctive-simpl} and the resolution
theorem of the \resolve rule.

>>>>

\begin{figure}
\centering
\begin{bprooftree}\tt
  \AxiomC{}
  \RightLabel{negate}
  \UnaryInfC{$\neg p$}
  \RightLabel{strip}
  \UnaryInfC{$\neg p$}
  \AxiomC{}
  \RightLabel{axiom}
  \UnaryInfC{$p$}
  \RightLabel{canonicalize}
  \UnaryInfC{$p$}
  \RightLabel{simplify}
  \BinaryInfC{$⊥$}
  \RightLabel{canonicalize}
  \UnaryInfC{$⊥$}
\end{bprooftree}
\caption{The \Metis' proof tree for $p \vdash p$ from the
derivation in Fig.~\ref{fig:metis-proof-tstp}}
\label{fig:metis-example}
\end{figure}

@@

\textit{Splitting}. To prove a goal, \Metis splits the goal into
disjoint cases. This process produces a list of new subgoals, the
conjunction of these subgoals implies the goal. Then, a proof of the
goal becomes in smaller proofs, one refutation for each subgoal.
These subgoals are introduced in the \TSTP derivation with the \strip
inference rule. We refer the reader to see Theorem~\ref{thm:thm-strip}.

@@

\textit{Clausification.} The \clausify rule transforms a
propositional formula in clausal normal form or \CNF, a conjunction
of clauses. Where a clause is the disjunction of zero or more
literals and a literal is an atom (positive literal) or a negation
of an atom (negative literal). This conversion is not unique, and
\Metis performs this transformation using its criteria.
We refer the reader to see Theorem~\ref{thm:thm-clausify}.

\improvement{We should mention that it was indeed a problem for the
proof-reconstruction? and refer the section next?}

@@

\textit{Normalization.} \Metis uses the \canonicalize rule to
normalize a propositional formula that comes from an axiom or a
definition. It often converts each expression in one of its normal
form: the conjunctive normal form, the negative normal form or the
disjunctive normal form. Next, this rule simplifies the formula
with some definitions, applying recursively on the structure's
formula conjunctions and disjunctions simplifications to remove
tautologies (see Fig.~\ref{fig:conjunctive-disjunctive-simpl} for a
list of these theorems. We assume the commutative property for the
conjunction and disjunction connectives).
We refer the reader to see Section~\ref{sssec:normalization} and the
lemmas and theorems described there.

@@

\textit{Resolution.} The \resolve rule is the resolution theorem
showed in Fig.~\ref{fig:metis-inferences}. For its application, we
need the positive literal for resolution and two derivations to
deduce the \emph{resolvent}. The positive literal $p$ must occur in
the formula from the first derivation, and the negative literal must
occur in the formula from the second derivation (this follows the
same pattern in \Metis' logic core in Fig.~\ref{fig:metis-inferences}
for the resolve inference).
We refer the reader to see Theorem~\ref{thm:thm-resolve}.

@@

\textit{Splitting a conjunct.} The \conjunct rule extracts from a
conjunction one of its conjuncts; it is a generalization of the
projection rules for the conjunction connective as the
following \TSTP derivation shows.
We refer the reader to see Theorem~\ref{thm:thm-conjunct}.

@@

\textit{Negate.}
Each subgoal proof is a refutation, thereby each proof assumes
the negation of its subgoal. The \negate rule
assumes the negation of a given subgoal. This rules always
appears after applies the \strip inference to the goal.

@@

\textit{Simplification.}
The \simplify rule is an inference that performs simplification of
definitions and tautologies. This rule
transverses a list of formulas by applying different theorems (e.g.,
\emph{modus pones}, \emph{modus tollens}, or \emph{disjunctive
syllogism}) including the list of theorems used by \canonicalize in
Fig. \ref{fig:conjunctive-disjunctive-simpl} and the resolution
theorem of the \resolve rule.
We refer the reader to see Section~\ref{sssec:simplify}.
